Visualizing demographic evolution using geographically inconsistent census data

### 2019-12-10 paper/discussion.tex
merged the published version of the paper with the actual code

Removing the need for geographical harmonisation greatly reduces the amount of
work necessary to explore demographic data, but the method still requires
consistent variables across the years. Matching the variables can be trivial for
some aspects (Age), but challenging for others (Income). The divulged income
ranges vary over time and the actual values change due to inflation. Some
variables were not considered in earlier censuses, such as Race in Canada, or
Hispanic population in the USA, hampering its use when they are available. Since
this is only a prototype, we matched few aspects, but a proper demographic
analysis would benefit from all available information.

>>>>

We removed the need for geographical harmonisation, but the method still
requires consistent variables across the years. Matching the variables can be
trivial for some aspects (Age), but challenging for others (Income). The
divulged income ranges vary over time and the actual values change due to
inflation. Since this is only a prototype, we matched few aspects, but a proper
demographic analysis would benefit from all available information.

### 2019-12-10 paper/method.tex
merged the published version of the paper with the actual code

Each area of each census year is represented as a node, and edges are placed
between nodes if the corresponding regions share geographic borders in the same
year. Further, edges are placed between nodes if the corresponding regions
belong to sequential years and their geometries intersect.
To avoid spurious connections caused by small fluctuations, one of the
geometries is slightly shrunk before the intersection, using a buffer of -1e-6
degrees. While the result would not change considerably by adding more
connections between nearby regions, these extra connections would greatly
increase the computational cost. More importantly, while weights will be
associated with these edges before they are processed, they are not derived from
the geometry, but from the data. The actual intersection area is not considered
in this representation.  This approach leads to a single network representing
the whole spatio-temporal space of the data.

@@

From network theory, a \emph{path} is a sequence of nodes of the network that
are connected pairwise by edges. As each node has associated temporal
information, we call \emph{temporal} any path such that the temporal information
increases along it. For instance, in Figure~\ref{fig:intuition}, the paths ACF,
ADF, ADG, and BEH are temporal paths, but not HEB, CDE, nor FDA. With harmonised
data, as illustrated by the path BEH, the time-series of each region would form
a temporal path, each node would be connected only to its older and newer
versions, and would belong to only one temporal path. Since our data is not
harmonised, more connections are allowed and each node can belong to an
arbitrary number of paths.

>>>>

Each area of each census year is represented as a node, and edges are placed
between nodes if the corresponding regions share geographic borders in the same
year. Further, edges are placed between nodes if the corresponding regions
belong to sequential years and their geometries intersect. To avoid spurious
connections caused by small fluctuations, one of the geometries is slightly
shrunk before the intersection, using a buffer of -1e-6 degrees, or about 10cm
at the equator. As long as this buffer is not large enough to remove significant
overlaps, its exact value will not have a significant impact on the results,
since they also rely on the data similarity, and it could safely be set to zero.
However, this would lead to extra connections that would increase the
computational cost. More importantly, while weights will be associated with
these edges before they are processed, they are not derived from the geometry,
but from the data. The actual intersection area is not considered in this
representation.  This approach leads to a single network representing the whole
spatio-temporal space of the data.

@@

From network theory, a \emph{path} is a sequence of nodes of the network that
are connected pairwise by edges. As each node has associated temporal
information, we call \emph{temporal} any path such that the temporal information
increases along it, representing information advancing in time. For instance, in
Figure~\ref{fig:intuition}, the paths ACF, ADF, ADG, and BEH are temporal paths,
but not HEB, CDE, nor FDA. With harmonised data, as illustrated by the path BEH,
the time-series of each region would form a temporal path, each node would be
connected only to its older and newer versions, and would belong to only one
temporal path. Since our data is not harmonised, more connections are allowed
and each node can belong to an arbitrary number of paths.
